---
title: "badger_MBIOL_report.Rmd"
output: html_document
date: "2026-03-13"
---
# Introduction to document
This Markdown file is a complete report for my MBiol ecological modelling project, to give a complete overview of the methodology to accompany write up. 
Sections cover:     
* Data loading and figure digitisation from literature
* Custom function creation  
* Model projection
* Analysis of model outputs

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## section 1 -  loading libraries
```{r}
library(ggplot2)
library(gridExtra)
library(tidyr)
library(tidyverse)
library(dplyr)
library(readr)
library(here)
```

Sourcing other scripts
```{r}
# # sourcing functions used throughout 
source(here("Code/Functions/01_all_functions.R"))

# data extraction
source(here("Code/02_data_extraction.R"))  # select options 2 (existing data) and 1 (all data)

# parameter definition and df set ups
source(here("Code/03_parameter_rate_setup.R"))

# setting up model scenarios
source(here("Code/04_model_projections.R"))  
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

